Coherence between subjective feeling and facial expression: A comparison of human and automatic recognition

### 2019-04-08 ACII_2019_paper/ACII_2019_paper.tex
final version

% Pandoc header
\usepackage{booktabs}
\usepackage{float}
\usepackage{tabu}

@@

Detractors of the Basic Emotion Theory consider emotion not as
genetically hardwired but as a learnt association between a given
situation and an appropriate response (Averill 1980; Barrett 2017a). For
the tenants of the constructivist approach, emotions are ``concepts''
based on past experiences and which are ``a collection of embodied,
whole brain representations that predict what is about to happen in the
sensory environment, what the best action is to deal with impending
events, and their consequences for allostasis'' (Barrett 2017b, p12).
Following this assumption, faces are best conceived as tools displaying
signals in social interactions (Crivelli and Fridlund 2018). These
signals can convey individuals' motivations and readiness (Frijda and
Tcherkassof 1997) or social messages (Fridlund and Rosenberg 1995).
Therefore, facial expressions are thought as behaviors which meaning is
inferred by the observer. Findings support this observer dependence
(Lindquist and Gendron 2013; Niedenthal et al. 2017). They show that to
make meaning of another person's facial behavior, the perceiver relies
in particular on her/his knowledge about emotion categories.

@@

\hypertarget{correlation-between-self-report-and-human-facial-expression-recognition}{%
\subsection{Correlation between self-report and human facial expression
recognition}\label{correlation-between-self-report-and-human-facial-expression-recognition}}

@@

\caption{\label{tab:confusionTable_sr_hr}\label{table:confusionTable_sr_hr}Agreement accuracy metrics for each emotion.}
\centering
\fontsize{8}{10}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
Emotion & Sensitivity & Specificity & Precision & F1\\
\midrule
anger & 0.43 & 0.93 & 0.23 & 0.3\\
disgust & 1.00 & 0.94 & 0.33 & 0.5\\
fear & 0.00 & 0.97 & 0.00 & \textit{na.}\\
happiness & 0.49 & 0.78 & 0.69 & 0.57\\
sadness & 0.08 & 0.92 & 0.08 & 0.08\\
surprise & 0.44 & 0.67 & 0.37 & 0.4\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} Note. \textit{na.} values are produced when not enough}\\
\multicolumn{5}{l}{data are available to compute accuracy indicators.}\\
\end{tabu}
\end{table}

@@

\hypertarget{correlation-between-self-report-and-automatic-facial-expression-recognition}{%
\subsection{Correlation between self-report and automatic facial
expression
recognition}\label{correlation-between-self-report-and-automatic-facial-expression-recognition}}

@@

\caption{\label{tab:confusionTable_sr_ar}\label{table:confusionTable_sr_ar}Agreement accuracy metrics for each emotion.}
\centering
\fontsize{8}{10}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
Emotion & Sensitivity & Specificity & Precision & F1\\
\midrule
anger & 0.00 & 0.90 & 0.00 & \textit{na.}\\
disgust & 0.25 & 0.72 & 0.02 & 0.04\\
fear & 0.00 & 0.94 & 0.00 & \textit{na.}\\
happiness & 0.44 & 0.73 & 0.62 & 0.52\\
sadness & 0.15 & 0.95 & 0.25 & 0.19\\
surprise & 0.19 & 0.86 & 0.36 & 0.25\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} Note. \textit{na.} values are produced when not enough}\\
\multicolumn{5}{l}{data are available to compute accuracy indicators.}\\
\end{tabu}
\end{table}

@@

The overall agreement between human observers and the automatic
classifier is in fact very low (Kappa = 0.18). Except for
\emph{happiness} and \emph{disgust} (respectively 64.9\% and 50.0\% of
common labelling), there is no clear common pattern. Moreover, the
automatic classifier has a tendency to label as \emph{disgust} videos
labeled as \emph{sadness} by human observers and as \emph{happiness}
videos labaled as \emph{fear} by human observers

@@

Some limitations should be stated, notably regarding the use of
self-reports to evaluate encoders' subjective feelings. Accessing the
inner subjective feeling can be biased if not impossible. Moreover the
procedure used for human observation can also be open to dispute.
Instead of asking the human annotators to provide a unique label, a more
subtle approach was chosen to mimic results provided by the automatic
classifier . Whereas this paradigm is longer and more complicated, it
can lead to more robust results in reducing the forced-choice biais
(Russell 1993). However this procedure can also reduce the human
observers accuracy. In this regard, the results of the human observation
could have been more ambiguous because it is not the natural way that
people are inferring meaning from facial expressions. An alternative
explanation relies in reducing the recognition biais involved in the
classic recognition paradigm. Classic forced-choice paradigms obtain
artificial high results, thus by using a more evolved approach
observers' accuracy may have been lowered.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-averill1980constructivist}{}%
Averill, James R. 1980. ``A Constructivist View of Emotion.'' In
\emph{Theories of Emotion}, 305--39. Elsevier.

\leavevmode\hypertarget{ref-baltruvsaitis2016openface}{}%
Baltrušaitis, Tadas, Peter Robinson, and Louis-Philippe Morency. 2016.
``Openface: An Open Source Facial Behavior Analysis Toolkit.'' In
\emph{2016 Ieee Winter Conference on Applications of Computer Vision
(Wacv)}, 1--10. IEEE.

\leavevmode\hypertarget{ref-barrett2017emotions}{}%
Barrett, Lisa Feldman. 2017a. \emph{How Emotions Are Made: The Secret
Life of the Brain}. Houghton Mifflin Harcourt.

\leavevmode\hypertarget{ref-barrett2017theory}{}%
---------. 2017b. ``The Theory of Constructed Emotion: An Active
Inference Account of Interoception and Categorization.'' \emph{Social
Cognitive and Affective Neuroscience} 12 (1): 1--23.

\leavevmode\hypertarget{ref-bonanno2004brief}{}%
Bonanno, George, and Dacher Keltner. 2004. ``Brief Report the Coherence
of Emotion Systems: Comparing `on-Line' Measures of Appraisal and Facial
Expressions, and Self-Report.'' \emph{Cognition and Emotion} 18 (3):
431--44.

\leavevmode\hypertarget{ref-crivelli2018facial}{}%
Crivelli, Carlos, and Alan J Fridlund. 2018. ``Facial Displays Are Tools
for Social Influence.'' \emph{Trends in Cognitive Sciences} 22 (5):
388--99.

\leavevmode\hypertarget{ref-darwin1872expression}{}%
Darwin, Charles. 1872. \emph{The Expression of the Emotions in Man and
Animals}. London, UK: John Murray, London UK.

\leavevmode\hypertarget{ref-dente2017measures}{}%
Dente, Pasquale, Dennis Küster, Lina Skora, and E Krumhuber. 2017.
``Measures and Metrics for Automatic Emotion Classification via Facet.''
In \emph{Proceedings of the Conference on the Study of Artificial
Intelligence and Simulation of Behaviour (Aisb)}, 160--63.

\leavevmode\hypertarget{ref-doyle2017language}{}%
Doyle, Cameron M, and Kristen A Lindquist. 2017. ``Language and Emotion:
Hypotheses on the Constructed Nature of Emotion Perception.''

\leavevmode\hypertarget{ref-dupre2015oudjat}{}%
Dupré, Damien, Daniel Akpan, Elena Elias, Jean-Michel Adam, Brigitte
Meillon, Nicolas Bonnefond, Michel Dubois, and Anna Tcherkassof. 2015.
``Oudjat: A Configurable and Usable Annotation Tool for the Study of
Facial Expressions of Emotion.'' \emph{International Journal of
Human-Computer Studies} 83: 51--61.

\leavevmode\hypertarget{ref-dupre2018accuracy}{}%
Dupré, Damien, Nicole Andelic, Gawain Morrison, and Gary McKeown. 2018.
``Accuracy of Three Commercial Automatic Emotion Recognition Systems
Across Different Individuals and Their Facial Expressions.'' In
\emph{2018 Ieee International Conference on Pervasive Computing and
Communications Workshops (Percom Workshops)}, 627--32. IEEE.

\leavevmode\hypertarget{ref-duran2017coherence}{}%
Durán, Juan I, Rainer Reisenzein, and José-Miguel Fernández-Dols. 2017.
``Coherence Between Emotions and Facial Expressions.'' \emph{The Science
of Facial Expression}, 107--29.

\leavevmode\hypertarget{ref-ekman1992argument}{}%
Ekman, Paul. 1992. ``An Argument for Basic Emotions.'' \emph{Cognition
\& Emotion} 6 (3-4): 169--200.

\leavevmode\hypertarget{ref-ekman2007directed}{}%
---------. 2007. ``The Directed Facial Action Task.'' \emph{Handbook of
Emotion Elicitation and Assessment} 47: 53.

\leavevmode\hypertarget{ref-ekman1987universals}{}%
Ekman, Paul, Wallace V Friesen, Maureen O'sullivan, Anthony Chan, Irene
Diacoyanni-Tarlatzis, Karl Heider, Rainer Krause, et al. 1987.
``Universals and Cultural Differences in the Judgments of Facial
Expressions of Emotion.'' \emph{Journal of Personality and Social
Psychology} 53 (4): 712.

\leavevmode\hypertarget{ref-ekman1988universality}{}%
Ekman, Paul, and Karl G Heider. 1988. ``The Universality of a Contempt
Expression: A Replication.'' \emph{Motivation and Emotion} 12 (3):
303--8.

\leavevmode\hypertarget{ref-fernandez2013emotion}{}%
Fernández-Dols, José-Miguel, and Carlos Crivelli. 2013. ``Emotion and
Expression: Naturalistic Studies.'' \emph{Emotion Review} 5 (1): 24--29.

\leavevmode\hypertarget{ref-fridlund1995human}{}%
Fridlund, Alan J, and Erika L Rosenberg. 1995. ``Human Facial
Expression: An Evolutionary View.'' \emph{Nature} 373 (6515): 569--69.

\leavevmode\hypertarget{ref-frijda1997facial}{}%
Frijda, Nico H, and Anna Tcherkassof. 1997. ``Facial Expressions as
Modes of Action Readiness.'' \emph{The Psychology of Facial Expression},
78--102.

\leavevmode\hypertarget{ref-kappas2003facial}{}%
Kappas, Arvid. 2003. ``What Facial Activity Can and Cannot Tell Us About
Emotions.'' In \emph{The Human Face}, 215--34. Springer.

\leavevmode\hypertarget{ref-kraut1979social}{}%
Kraut, Robert E, and Robert E Johnston. 1979. ``Social and Emotional
Messages of Smiling: An Ethological Approach.'' \emph{Journal of
Personality and Social Psychology} 37 (9): 1539.

\leavevmode\hypertarget{ref-lindquist2013s}{}%
Lindquist, Kristen A, and Maria Gendron. 2013. ``What's in a Word?
Language Constructs Emotion Perception.'' \emph{Emotion Review} 5 (1):
66--71.

\leavevmode\hypertarget{ref-mcduff2016discovering}{}%
McDuff, Daniel. 2016. ``Discovering Facial Expressions for States of
Amused, Persuaded, Informed, Sentimental and Inspired.'' In
\emph{Proceedings of the 18th Acm International Conference on Multimodal
Interaction}, 71--75. ACM.

\leavevmode\hypertarget{ref-mcduff2016affdex}{}%
McDuff, Daniel, Abdelrahman Mahmoud, Mohammad Mavadati, May Amr, Jay
Turcot, and Rana el Kaliouby. 2016. ``AFFDEX Sdk: A Cross-Platform
Real-Time Multi-Face Expression Recognition Toolkit.'' In
\emph{Proceedings of the 2016 Chi Conference Extended Abstracts on Human
Factors in Computing Systems}, 3723--6. ACM.

\leavevmode\hypertarget{ref-mckeown2013analogical}{}%
McKeown, Gary J. 2013. ``The Analogical Peacock Hypothesis: The Sexual
Selection of Mind-Reading and Relational Cognition in Human
Communication.'' \emph{Review of General Psychology} 17 (3): 267--87.

\leavevmode\hypertarget{ref-niedenthal2017embodied}{}%
Niedenthal, Paula M, Adrienne Wood, Magdalena Rychlowska, and Sebastian
Korb. 2017. ``Embodied Simulation in Decoding Facial Expression.''
\emph{The Science of Facial Expression}, 397.

\leavevmode\hypertarget{ref-russell1993forced}{}%
Russell, James A. 1993. ``Forced-Choice Response Format in the Study of
Facial Expression.'' \emph{Motivation and Emotion} 17 (1): 41--51.

\leavevmode\hypertarget{ref-tcherkassof2013dynemo}{}%
Tcherkassof, Anna, Damien Dupré, Brigitte Meillon, Nadine Mandran,
Michel Dubois, and Jean-Michel Adam. 2013. ``DynEmo: A Video Database of
Natural Facial Expressions of Emotions.'' \emph{The International
Journal of Multimedia \& Its Applications} 5 (5): 61--80.

\leavevmode\hypertarget{ref-de2019mama}{}%
Waal, Frans de. 2019. \emph{Mama's Last Hug: Animal Emotions and What
They Teach Us About Ourselves}. Granta Books.

>>>>

% Pandoc header
\usepackage{booktabs}
\usepackage{float}
\usepackage{tabu}
\usepackage{wrapfig}
\floatplacement{figure}{H}
\floatplacement{table}{H}

@@

Detractors of the Basic Emotion Theory consider emotion not as
genetically hardwired but as a learnt association between a given
situation and an appropriate response (Averill 1980; Barrett 2017a). For
the tenants of the constructivist approach, emotions are ``concepts''
based on past experiences and which are ``a collection of embodied,
whole brain representations that predict what is about to happen in the
sensory environment, what the best action is to deal with impending
events, and their consequences for allostasis'' (Barrett 2017b, p12).
Following this assumption, faces are best conceived as tools displaying
signals in social interactions (Crivelli and Fridlund 2018). These
signals can convey individuals' motivations and readiness (Frijda and
Tcherkassof 1997) or social messages (Fridlund 1994). Therefore, facial
expressions are thought as behaviors which meaning is inferred by the
observer. Findings support this observer dependence (Lindquist and
Gendron 2013; Niedenthal et al. 2017). They show that to make meaning of
another person's facial behavior, the perceiver relies in particular on
her/his knowledge about emotion categories.

@@

\hypertarget{self-reported-emotions-recognized-by-human-observers}{%
\subsection{Self-reported emotions recognized by human
observers}\label{self-reported-emotions-recognized-by-human-observers}}

@@

\begin{table}[H]

\caption{\label{tab:confusionTable_sr_hr}\label{table:confusionTable_sr_hr}Human recognition accuracy metrics for each emotion.}
\centering
\fontsize{8}{10}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
Emotion & Sensitivity & Specificity & Precision & F1\\
\midrule
anger & 0.43 & 0.93 & 0.23 & 0.3\\
disgust & 1.00 & 0.94 & 0.33 & 0.5\\
fear & 0.00 & 0.97 & 0.00 & \textit{na.}\\
happiness & 0.49 & 0.78 & 0.69 & 0.57\\
sadness & 0.08 & 0.92 & 0.08 & 0.08\\
surprise & 0.44 & 0.67 & 0.37 & 0.4\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} Note. \textit{na.} values are produced when not enough}\\
\multicolumn{5}{l}{data are available to compute accuracy indicators.}\\
\end{tabu}
\end{table}

@@

\hypertarget{self-reported-emotions-recognized-by-the-automatic-classifier}{%
\subsection{Self-reported emotions recognized by the automatic
classifier}\label{self-reported-emotions-recognized-by-the-automatic-classifier}}

@@

\caption{\label{tab:confusionTable_sr_ar}\label{table:confusionTable_sr_ar}Autonatic recognition accuracy metrics for each emotion.}
\centering
\fontsize{8}{10}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
Emotion & Sensitivity & Specificity & Precision & F1\\
\midrule
anger & 0.00 & 0.90 & 0.00 & \textit{na.}\\
disgust & 0.25 & 0.72 & 0.02 & 0.04\\
fear & 0.00 & 0.94 & 0.00 & \textit{na.}\\
happiness & 0.44 & 0.73 & 0.62 & 0.52\\
sadness & 0.15 & 0.95 & 0.25 & 0.19\\
surprise & 0.19 & 0.86 & 0.36 & 0.25\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} Note. \textit{na.} values are produced when not enough}\\
\multicolumn{5}{l}{data are available to compute accuracy indicators.}\\
\end{tabu}
\end{table}

@@

The overall agreement between human observers and the automatic
classifier is in fact very low (Kappa = 0.18). Except for
\emph{happiness} and \emph{disgust} (respectively 64.9\% and 50.0\% of
common labelling), there is no clear common pattern. Moreover, the
automatic classifier has a tendency to label as \emph{disgust} videos
labeled as \emph{sadness} by human observers and as \emph{happiness}
videos labeled as \emph{fear} by human observers

@@

Some limitations should be stated, notably regarding the use of
self-reports to evaluate encoders' subjective feelings. Accessing the
inner subjective feeling can be biased if not impossible. Moreover the
procedure used for human observation can also be open to dispute.
Instead of asking the human annotators to provide a unique label, a more
subtle approach was chosen to mimic results provided by the automatic
classifier . Whereas this paradigm is longer and more complicated, it
can lead to more robust results in reducing the forced-choice bais
(Russell 1993). However this procedure can also reduce the human
observers accuracy. In this regard, the results of the human observation
could have been more ambiguous because it is not the natural way that
people are inferring meaning from facial expressions. An alternative
explanation relies in reducing the recognition bais involved in the
classic recognition paradigm. Classic forced-choice paradigms obtain
artificial high results, thus by using a more evolved approach
observers' accuracy may have been lowered.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-averill1980constructivist}{}%
Averill, James R. 1980. ``A Constructivist View of Emotion.'' In
\emph{Theories of Emotion}, edited by Robert Plutchik and Henry
Kellerman, 305--39. New York, NY: Academic Press.

\leavevmode\hypertarget{ref-baltruvsaitis2016openface}{}%
Baltrušaitis, Tadas, Peter Robinson, and Louis-Philippe Morency. 2016.
``Openface: An Open Source Facial Behavior Analysis Toolkit.'' In
\emph{Proceedings of the Winter Conference on Applications of Computer
Vision}, 1--10. \url{https://doi.org/10.1109/WACV.2016.7477553}.

\leavevmode\hypertarget{ref-barrett2017emotions}{}%
Barrett, Lisa Feldman. 2017a. \emph{How Emotions Are Made: The Secret
Life of the Brain}. Boston, MA: Houghton Mifflin Harcourt.

\leavevmode\hypertarget{ref-barrett2017theory}{}%
---------. 2017b. ``The Theory of Constructed Emotion: An Active
Inference Account of Interoception and Categorization.'' \emph{Social
Cognitive and Affective Neuroscience} 12 (1): 1--23.
\url{https://doi.org/10.1093/scan/nsw154}.

\leavevmode\hypertarget{ref-bonanno2004brief}{}%
Bonanno, George, and Dacher Keltner. 2004. ``The Coherence of Emotion
Systems: Comparing `on-Line' Measures of Appraisal and Facial
Expressions, and Self-Report.'' \emph{Cognition and Emotion} 18 (3):
431--44. \url{https://doi.org/10.1080/02699930341000149}.

\leavevmode\hypertarget{ref-crivelli2018facial}{}%
Crivelli, Carlos, and Alan J Fridlund. 2018. ``Facial Displays Are Tools
for Social Influence.'' \emph{Trends in Cognitive Sciences} 22 (5):
388--99. \url{https://doi.org/10.1016/j.tics.2018.02.006}.

\leavevmode\hypertarget{ref-darwin1872expression}{}%
Darwin, Charles. 1872. \emph{The Expression of the Emotions in Man and
Animals}. London, UK: John Murray.

\leavevmode\hypertarget{ref-dente2017measures}{}%
Dente, Pasquale, Dennis Küster, Lina Skora, and E Krumhuber. 2017.
``Measures and Metrics for Automatic Emotion Classification via Facet.''
In \emph{Proceedings of the Conference on the Study of Artificial
Intelligence and Simulation of Behaviour}, 160--63.

\leavevmode\hypertarget{ref-doyle2017language}{}%
Doyle, Cameron M, and Kristen A Lindquist. 2017. ``Language and Emotion:
Hypotheses on the Constructed Nature of Emotion Perception.'' In
\emph{The Science of Facial Expression}, edited by José-Miguel
Fernández-Dols and James A Russell, 415--32. New York, NY: Oxford
University Press.

\leavevmode\hypertarget{ref-dupre2015oudjat}{}%
Dupré, Damien, Daniel Akpan, Elena Elias, Jean-Michel Adam, Brigitte
Meillon, Nicolas Bonnefond, Michel Dubois, and Anna Tcherkassof. 2015.
``Oudjat: A Configurable and Usable Annotation Tool for the Study of
Facial Expressions of Emotion.'' \emph{International Journal of
Human-Computer Studies} 83: 51--61.
\url{https://doi.org/10.1016/j.ijhcs.2015.05.010}.

\leavevmode\hypertarget{ref-dupre2018accuracy}{}%
Dupré, Damien, Nicole Andelic, Gawain Morrison, and Gary McKeown. 2018.
``Accuracy of Three Commercial Automatic Emotion Recognition Systems
Across Different Individuals and Their Facial Expressions.'' In
\emph{Proceedings of the International Conference on Pervasive Computing
and Communications}, 627--32.
\url{https://doi.org/10.1109/PERCOMW.2018.8480127}.

\leavevmode\hypertarget{ref-duran2017coherence}{}%
Durán, Juan I, Rainer Reisenzein, and José-Miguel Fernández-Dols. 2017.
``Coherence Between Emotions and Facial Expressions.'' Edited by
José-Miguel Fernández-Dols and James A Russell, 107--29.

\leavevmode\hypertarget{ref-ekman1992argument}{}%
Ekman, Paul. 1992. ``An Argument for Basic Emotions.'' \emph{Cognition
\& Emotion} 6 (3-4): 169--200.
\url{https://doi.org/10.1080/02699939208411068}.

\leavevmode\hypertarget{ref-ekman2007directed}{}%
---------. 2007. ``The Directed Facial Action Task.'' In \emph{Handbook
of Emotion Elicitation and Assessment}, edited by James A. Coan and John
J. B. Allen, 47--53. New York, NY: Oxford University Press.

\leavevmode\hypertarget{ref-ekman1987universals}{}%
Ekman, Paul, Wallace V Friesen, Maureen O'sullivan, Anthony Chan, Irene
Diacoyanni-Tarlatzis, Karl Heider, Rainer Krause, et al. 1987.
``Universals and Cultural Differences in the Judgments of Facial
Expressions of Emotion.'' \emph{Journal of Personality and Social
Psychology} 53 (4): 712--17.
\url{https://doi.org/10.1037/0022-3514.53.4.712}.

\leavevmode\hypertarget{ref-ekman1988universality}{}%
Ekman, Paul, and Karl G Heider. 1988. ``The Universality of a Contempt
Expression: A Replication.'' \emph{Motivation and Emotion} 12 (3):
303--8. \url{https://doi.org/10.1007/BF00993116}.

\leavevmode\hypertarget{ref-fernandez2013emotion}{}%
Fernández-Dols, José-Miguel, and Carlos Crivelli. 2013. ``Emotion and
Expression: Naturalistic Studies.'' \emph{Emotion Review} 5 (1): 24--29.
\url{https://doi.org/10.1177/1754073912457229}.

\leavevmode\hypertarget{ref-fridlund1994human}{}%
Fridlund, Alan J. 1994. \emph{Human Facial Expression: An Evolutionary
View}. San Diego, CA: Academic Press.

\leavevmode\hypertarget{ref-frijda1997facial}{}%
Frijda, Nico H, and Anna Tcherkassof. 1997. ``Facial Expressions as
Modes of Action Readiness.'' In \emph{The Psychology of Facial
Expression}, edited by James A Russell and José-Miguel Fernández-Dols,
78--102. Cambridge, UK: Cambridge University Press.

\leavevmode\hypertarget{ref-kappas2003facial}{}%
Kappas, Arvid. 2003. ``What Facial Activity Can and Cannot Tell Us About
Emotions.'' In \emph{The Human Face}, edited by Mary Katsikitis,
215--34. Boston, MA: Springer.

\leavevmode\hypertarget{ref-kraut1979social}{}%
Kraut, Robert E, and Robert E Johnston. 1979. ``Social and Emotional
Messages of Smiling: An Ethological Approach.'' \emph{Journal of
Personality and Social Psychology} 37 (9): 1539--53.

\leavevmode\hypertarget{ref-lindquist2013s}{}%
Lindquist, Kristen A, and Maria Gendron. 2013. ``What's in a Word?
Language Constructs Emotion Perception.'' \emph{Emotion Review} 5 (1):
66--71. \url{https://doi.org/10.1177/1754073912451351}.

\leavevmode\hypertarget{ref-mcduff2016discovering}{}%
McDuff, Daniel. 2016. ``Discovering Facial Expressions for States of
Amused, Persuaded, Informed, Sentimental and Inspired.'' In
\emph{Proceedings of the International Conference on Multimodal
Interaction}, 71--75. \url{https://doi.org/10.1145/2993148.2993192}.

\leavevmode\hypertarget{ref-mcduff2016affdex}{}%
McDuff, Daniel, Abdelrahman Mahmoud, Mohammad Mavadati, May Amr, Jay
Turcot, and Rana el Kaliouby. 2016. ``AFFDEX Sdk: A Cross-Platform
Real-Time Multi-Face Expression Recognition Toolkit.'' In
\emph{Proceedings of the Chi Conference on Human Factors in Computing
Systems}, 3723--6. \url{https://doi.org/10.1145/2851581.2890247}.

\leavevmode\hypertarget{ref-mckeown2013analogical}{}%
McKeown, Gary J. 2013. ``The Analogical Peacock Hypothesis: The Sexual
Selection of Mind-Reading and Relational Cognition in Human
Communication.'' \emph{Review of General Psychology} 17 (3): 267--87.
\url{https://doi.org/10.1037/a0032631}.

\leavevmode\hypertarget{ref-niedenthal2017embodied}{}%
Niedenthal, Paula M, Adrienne Wood, Magdalena Rychlowska, and Sebastian
Korb. 2017. ``Embodied Simulation in Decoding Facial Expression.'' In
\emph{The Science of Facial Expression}, edited by José-Miguel
Fernández-Dols and James A Russell, 397--414. New York, NY: Oxford
University Press.

\leavevmode\hypertarget{ref-russell1993forced}{}%
Russell, James A. 1993. ``Forced-Choice Response Format in the Study of
Facial Expression.'' \emph{Motivation and Emotion} 17 (1): 41--51.
\url{https://doi.org/10.1007/BF00995206}.

\leavevmode\hypertarget{ref-tcherkassof2013dynemo}{}%
Tcherkassof, Anna, Damien Dupré, Brigitte Meillon, Nadine Mandran,
Michel Dubois, and Jean-Michel Adam. 2013. ``DynEmo: A Video Database of
Natural Facial Expressions of Emotions.'' \emph{The International
Journal of Multimedia \& Its Applications} 5 (5): 61--80.
\url{https://doi.org/10.5121/ijma.2013.5505}.

\leavevmode\hypertarget{ref-de2019mama}{}%
Waal, Frans de. 2019. \emph{Mama's Last Hug: Animal Emotions and What
They Teach Us About Ourselves}. London, UK: Granta Books.
